---
title: "TO404HW2"
author: "Maxwell Cornellier"
date: "11/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Library Inputs 

```{r}

library(dplyr)
library("ggplot2")
```

# 1. Understand the Data 


###Download Data Files & Take 5% Sample of Each 

```{r}
Jan2019 <- read.csv("201901-citibike-tripdata.csv")
SampleJan2019 <- sample_frac(Jan2019,0.05)
SampleFeb2019 <- sample_frac((read.csv("201902-citibike-tripdata.csv")),0.05)
SampleMar2019 <- sample_frac((read.csv("201903-citibike-tripdata.csv")),0.05)
SampleApril2019 <- sample_frac((read.csv("201904-citibike-tripdata.csv")),0.05)
SampleMay2019 <- sample_frac((read.csv("201905-citibike-tripdata.csv")),0.05)
SampleJune2019 <- sample_frac((read.csv("201906-citibike-tripdata.csv")),0.05)
SampleJuly2019 <- sample_frac((read.csv("201907-citibike-tripdata.csv")),0.05)
SampleAugust2019 <- sample_frac((read.csv("201908-citibike-tripdata.csv")),0.05)
SampleSept2019 <- sample_frac((read.csv("201909-citibike-tripdata.csv")),0.05)
SampleOct2019 <- sample_frac((read.csv("201910-citibike-tripdata.csv")),0.05)
SampleNov2019 <- sample_frac((read.csv("201911-citibike-tripdata.csv")),0.05) 
SampleDec2019 <- sample_frac((read.csv("201912-citibike-tripdata.csv")),0.05)
```

### Combine Data Files into one Table 

```{r}

CitiBike2019 <- rbind(SampleJan2019,SampleFeb2019,SampleMar2019, SampleApril2019, SampleMay2019, SampleJune2019, SampleJuly2019, SampleAugust2019, SampleSept2019, SampleOct2019, SampleNov2019, SampleDec2019)

# Clean Data 

## Change Variables to Factors as necessary 

CitiBike2019$start.station.id <- as.factor(CitiBike2019$start.station.id)
CitiBike2019$end.station.id <- as.factor(CitiBike2019$end.station.id)
CitiBike2019$end.station.name <- as.factor(CitiBike2019$end.station.name)
CitiBike2019$start.station.name <- as.factor(CitiBike2019$start.station.name)
CitiBike2019$usertype <- as.factor(CitiBike2019$usertype)
CitiBike2019$gender <- as.factor(CitiBike2019$gender)
CitiBike2019$bikeid <- as.factor(CitiBike2019$bikeid)
CitiBike2019$starttime <- as.POSIXct(CitiBike2019$starttime)
CitiBike2019$stoptime <- as.POSIXct(CitiBike2019$stoptime)
```


## Descriptive Stats and Visualizations 


```{r}
# Average Trip Duration 

arr_tripduration <- mean(CitiBike2019$tripduration, na.rm = T)

# Longest trip duration 

longest_trip<- max(CitiBike2019$tripduration, na.rm = T)


# Shortest trip duration 

shortest_trip <- min(CitiBike2019$tripduration, na.rm = T)
```
The average trip duration for a Citi Bike ride in 2019 was `r arr_tripduration` seconds 

The longest trip by duration on a Citi Bike in 2019 was `r longest_trip` seconds or about 38.68 Days 

The shortest trip by duration on a Citi Bike in 2019 was `r shortest_trip` seconds


#Stats, Visualizations, and Patterns 


```{r}
# Number of Citi Bike Stations 

number_endstations <- length(levels(CitiBike2019$end.station.id)) 

#Citi bike Rides in 2019 by User type 

CountCustumer <- nrow(CitiBike2019[CitiBike2019$usertype == "Customer",])
CountSubscriber <- nrow(CitiBike2019[CitiBike2019$usertype == "Subscriber",])
TotalRides <- nrow(CitiBike2019)

baseplot <- ggplot(data=CitiBike2019, aes(x= usertype))
baseplot + geom_histogram(stat="count")
```

Number of Citi Bike End Station is `r number_endstations`

Number of Citi Bike rides by users who are "Customers" is `r CountCustumer`
Number of Citi Bike rides by users who are "Subscribers" is `r CountSubscriber`

That means that 13.88% of rides are done by customers while 86.12% of rides are done by subscribers 


```{r}


#Average Trip Duration by User Type 

avgtrip_usertype <- tapply(CitiBike2019$tripduration,CitiBike2019$usertype, mean, na.rm = T)


barplot(sort(avgtrip_usertype, decreasing = T))
```

```{r}
# User Type by Age Group 




baseplot2 <- ggplot(data=CitiBike2019, aes( x= birth.year ), na.rm = T)
baseplot2 + geom_histogram()
```




```{r}

# Average User Age 

average_userage <- mean(CitiBike2019$birth.year)

# trip duration by month 

baseplot3 <- ggplot(data = CitiBike2019, aes(x= starttime), na.rm = T)
baseplot3 + geom_histogram()


```

The average year of birth of a Citi Bike user in 2019 is `r average_userage` making them about 39 years of age.

### Observations 

Some takeaways are that Citi Bike riders are mainly subscribers that have short trip duration compared to a small percentage of Citi Bike Users who are "Customers" who have longer trip durations on average 

It can also be seen that there are more start times during the summer months which makes sense given the weather in NYC and the limitations of bicycles in snowy conditions. 

Citi Bike users are on average around 39 years old and face lower numbers among younger customers. 






